### 2020-04-13 ms/__latexindent_temp.tex
submit for review

@inproceedings{van_strien_assessing_2020,
  title = {Assessing the {{Impact}} of {{OCR Quality}} on {{Downstream NLP Tasks}}:},
  shorttitle = {Assessing the {{Impact}} of {{OCR Quality}} on {{Downstream NLP Tasks}}},
  booktitle = {Proceedings of the 12th {{International Conference}} on {{Agents}} and {{Artificial Intelligence}}},
  author = {{van Strien}, Daniel and Beelen, Kaspar and Ardanuy, Mariona and Hosseini, Kasra and McGillivray, Barbara and Colavizza, Giovanni},
  year = {2020},
  pages = {484--496},
  publisher = {{SCITEPRESS - Science and Technology Publications}},
  address = {{Valletta, Malta}},
  doi = {10.5220/0009169004840496},
  file = {/Users/melvinwevers/Zotero/storage/FRZEFAYM/van Strien et al. - 2020 - Assessing the Impact of OCR Quality on Downstream .pdf},
  isbn = {978-989-758-395-7},
  language = {en}
}